---
title: "Change over time? Interest in politics"
format: html
Jupyter: ir
editor: visual
---

This book explores whether interest in politics has changed over time. Follow the code below, to examine data from the British Social Attitudes survey from 2017, 2019 and 2021.

```{r}
library(haven)  ### To open Stata or SPSS datasets
library(dplyr)
library(ggplot2)
library(srvyr)  # simplified syntax for survey analysis

```

## Data

The exercise uses three datasets from the British Social Attitudes Survey. Start by reading in the data for the three years.

```{r read data}
BSA2017 <- read_dta("https://github.com/UKDataServiceOpen/Open_data/raw/refs/heads/master/bsa2017_open_enviropol.dta")
BSA2019<- read_dta("https://github.com/UKDataServiceOpen/Open_data/raw/refs/heads/master/bsa2019_poverty_open.dta")
BSA2021<- read_dta("https://github.com/UKDataServiceOpen/Open_data/raw/refs/heads/master/bsa2021_healthcare_equalities_open.dta")



library(haven)

# Remove labels and ensure numeric
BSA2021 <- BSA2021 %>%
  mutate(BSA21_final_wt = zap_labels(BSA21_final_wt))  # removes attached labels

BSA2017 <- BSA2017 %>%
  mutate(WtFactor = zap_labels(WtFactor),
         WtFactor = ifelse(WtFactor < 0, NA, WtFactor))

BSA2019 <- BSA2019 %>%
  mutate(WtFactor = zap_labels(WtFactor),
         WtFactor = ifelse(WtFactor < 0, NA, WtFactor))



```

## Interest in politics

```{r}
BSA2017$Politics<-as_factor(BSA2017$Politics)      #Create a new factor variable 
BSA2019$Politics<-as_factor(BSA2019$Politics)  
BSA2021$Politics<-as_factor(BSA2021$Politicsx)  


BSA2017 |> select(Politics) |> 
           summary()
BSA2019 |> select(Politics) |> 
           summary()
BSA2021 |> select(Politics) |> 
           summary()




```

Eeek! our variable needs harmonizing across the years.

```{r}
harmonise_politics <- function(x) {
  x <- tolower(trimws(as.character(x)))
  case_when(
    x %in% c("... a great deal,", "a great deal", "quite a lot,", "quite a lot") ~ "Great deal or quite a lot",
    x %in% c("some,", "some") ~ "Some",
    x %in% c("not very much,", "not very much") ~ "Not very much",
    x %in% c("or, none at all?", "none at all") ~ "None at all",
    TRUE ~ NA_character_
  )
}


```

```{r}
BSA2017_clean <- BSA2017 %>%
  rename(weight = WtFactor) %>% #rename weight
  mutate(
    Politics = harmonise_politics(Politics), #harmonise Politics variable
    Year = 2017 #create variable to indicate year
  ) %>%
  filter(!is.na(Politics)) %>% # Remove cases where Politics is NA (e.g. refusals, don't knows)
  select(Politics, weight, Year) #Selecting variables we need

BSA2019_clean <- BSA2019 %>%
  rename(weight = WtFactor) %>%
  mutate(
    Politics = harmonise_politics(Politics),
    Year = 2019
  ) %>%
  filter(!is.na(Politics)) %>%
  select(Politics, weight, Year)

BSA2021_clean <- BSA2021 %>%
  rename(weight = BSA21_final_wt) %>%
  mutate(
    Politics = harmonise_politics(Politics),
    Year = 2021
  ) %>%
  filter(!is.na(Politics)) %>%
  select(Politics, weight, Year)

```

```{r}
# Combine datasets if you haven’t already
combined_data <- bind_rows(BSA2017_clean, BSA2019_clean, BSA2021_clean)

# Set up survey design
svy_data <- as_survey_design(combined_data, ids = NULL, weights = "weight")


# Calculate weighted proportions by year and interest category

politics_summary <- svy_data %>%
  filter(Politics == "Great deal or quite a lot") %>%
  group_by(Year) %>%
  summarise(proportion = survey_mean(vartype = "ci"), .groups = "drop")



```

```{r}
library(ggplot2)

ggplot(politics_summary, aes(x = Year, y = proportion)) +
  geom_line(size = 1.2, colour = "steelblue") +
  geom_point(size = 3, colour = "steelblue") +
  geom_errorbar(aes(ymin = proportion_low, ymax = proportion_upp), width = 0.2, linewideth=1) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Proportion Saying 'Great deal or quite a lot' of Interest in Politics",
    x = "Year",
    y = "Weighted Percentage"
  ) +
  theme_minimal()


```

```         
```

```         
```
